
\chapter{Coordination Avoidance and RAMP Transactions}
\label{c.ramp}

In the previous chapter, we identified several existing isolation and
distributed consistency guarantees as coordination-free. Our goal was
proof-of-concept algorithms and systems support for these levels. In
this section, we go further. First, we develop a \textit{new} isolation
model that is tailored to a set of existing use cases for which there
is no existing, sufficiently powerful \iconfluent semantics. Second,
we develop high performance algorithms for enforcing those
semantics. 

Specifically, in this chapter, we address a largely underserved class
of applications requiring multi-partition, atomically
visible\footnote{\label{atomicnote}Our use of ``atomic''
  (specifically, Read Atomic isolation) concerns all-or-nothing
  \textit{visibility} of updates (i.e., the ACID isolation effects of
  ACID atomicity; Section~\ref{sec:ra-def}). This differs from uses of
  ``atomicity'' to denote serializability~\cite{bernstein-book} or
  linearizability~\cite{dcomp-book}.}  \textit{transactional} access:
cases where all or none of each transaction's effects should be
visible. In fact, this access corresponds to two semantics from the
previous chapter: MAV combined with Cut Isolation. The status quo for
these multi-partition atomic transactions provides an uncomfortable
choice between algorithms that are fast but deliver inconsistent
results and algorithms that deliver consistent results but are often
slow and unavailable under failure. Many of the largest modern,
real-world systems opt for protocols that guarantee fast and scalable
operation but provide few---if any---transactional semantics for
operations on arbitrary sets of data
items~\cite{tao,bigtable,pnuts,dynamo,2pc-scalability,espresso,rainbird}. This
may lead to anomalous behavior for several use cases requiring atomic
visibility, including secondary indexing, foreign key constraint
enforcement, and materialized view maintenance
(Section~\ref{sec:motivation}). In contrast, many traditional
transactional mechanisms correctly ensure atomicity of
updates~\cite{bernstein-book,spanner,calvin}. However, these
algorithms---such as two-phase locking and variants of optimistic
concurrency control---are often coordination-intensive, slow, and,
under failure, unavailable in a distributed
environment~\cite{hat-vldb,schism,jones-dtxn,pavlo-partition}. This
specific dichotomy between scalability and atomic visibility has been
described as ``a fact of life in the big cruel world of huge
systems''~\cite{helland-trans}.

Our contribution in this chapter is to demonstrate that atomically
visible transactions on partitioned databases are \textit{not} at odds
with scalability. We provide high-performance implementations of a
new, non-serializable isolation model called Read Atomic (RA)
isolation, corresponding to MAV with Cut Isolation. RA ensures that
all or none of each transaction's updates are visible to others and
that each transaction reads from an atomic snapshot of database state
(Section~\ref{sec:ra-def})---this is useful in the applications we
target. We subsequently develop three new, scalable algorithms for
achieving RA isolation that we collectively title Read Atomic
Multi-Partition (RAMP) transactions
(Section~\ref{sec:algorithm}). RAMP transactions guarantee scalability
and outperform existing atomic algorithms because they satisfy two key
scalability constraints. First, RAMP transactions guarantee
coordination-free execution: per Chapter~\ref{c.background}, one
client's transactions cannot cause another client's transactions to
stall or fail. Second, RAMP transactions guarantee \textit{partition
  independence}: clients only contact partitions that their
transactions directly reference (i.e., there is no central master,
coordinator, or scheduler). Together, these properties ensure
guaranteed completion, limited coordination across partitions, and
horizontal scalability for multi-partition access.

RAMP transactions are scalable because they appropriately control the
visibility of updates without inhibiting concurrency. Rather than
force concurrent reads and writes to stall, RAMP transactions allow
reads to ``race'' writes: RAMP transactions can autonomously detect
the presence of non-atomic (partial) reads and, if necessary, repair
them via a second round of communication with servers. To accomplish
this, RAMP writers attach metadata to each write and use limited
multi-versioning to prevent readers from stalling. The three
algorithms we present offer a trade-off between the size of this
metadata and performance. \texttt{RAMP-Small} transactions require
constant space (a timestamp per write) and two round trip time delays
(RTTs) for reads and writes. \texttt{RAMP-Fast} transactions require
metadata size that is linear in the number of writes in the
transaction but only require one RTT for reads in the common case and
two in the worst case. \texttt{RAMP-Hybrid} transactions employ Bloom
filters~\cite{bloomfilter} to provide an intermediate
solution. Traditional techniques like locking couple atomic visibility
and mutual exclusion; RAMP transactions provide the benefits of the
former without incurring the scalability, availability, or latency
penalties of the latter.

In addition to providing a theoretical analysis and proofs of
correctness, we demonstrate that RAMP transactions deliver in
practice. Our RAMP implementation achieves linear scalability to over
$7$ million operations per second on a 100 server cluster (at overhead
below $5\%$ for a workload of $95\%$ reads). Moreover, across a range
of workload configurations, RAMP transactions incur limited overhead
compared to other techniques and achieve higher performance than
existing approaches to atomic visibility
(Section~\ref{sec:ramp-evaluation}).

While the literature contains an abundance of isolation models, we
believe that the large number of modern applications requiring RA
isolation and the excellent scalability of RAMP transactions justify
the addition of yet another model. RA isolation is too weak for some
applications, but, for the many that it can serve, RAMP transactions
offer substantial benefits.

The remainder of this article proceeds as follows:
Section~\ref{sec:ramp-motivation} presents an overview of RAMP
transactions and describes key use cases based on industry
reports. Section~\ref{sec:ra-def} defines Read Atomic isolation,
presents both a detailed comparison with existing isolation guarantees
and a syntactic condition, the Read-Subset-Writes property, that
guarantees equivalence to serializable isolation, and defines two key
scalability criteria for RAMP algorithms to
provide. Section~\ref{sec:ramp-algorithm} presents and analyzes three
RAMP algorithms, which we experimentally evaluate in
Section~\ref{sec:ramp-evaluation}. Section~\ref{sec:ramp-application}
presents modifications of the RAMP protocols to better support
multi-datacenter deployments and to enforce transitive
dependencies. Section~\ref{sec:conclusion} concludes with a discussion
of extensions to the protocols presented here.


\input{ramp/motivation}

\input{ramp/model}

\input{ramp/algorithm}

\input{ramp/evaluation}

\input{ramp/application}

\input{ramp/relatedwork}

\input{ramp/appendix}

\input{ramp/conclusion}

